MSDS 6371 Project - Fall 2022

### 2022-12-08 Analysis 2.qmd
Writing updates

Forward step regression using olss
```{r}
forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
forwardModel = forwardResults$model
summary(forwardModel)

plot(forwardResults)

plot(forwardModel)
#performance::check_model(forwardModel, check = c("linearity", "normality", "qq"))

results = housesTest
results$SalePrice = exp(predict.lm(forwardModel, newdata = housesTest))
results = results |>
  dplyr::select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Forward_Model.csv")
```

@@

# backwardModel = lm(SalePrice ~ LotArea + Street + LotConfig + LandSlope + Neighborhood + Condition1 + Condition2 + BldgType + OverallQual + OverallCond + YearBuilt + YearRemodAdd + RoofMatl + ExterQual + ExterCond + Foundation + Heating + HeatingQC + CentralAir + `1stFlrSF` + `2ndFlrSF` + GrLivArea + HalfBath + BedroomAbvGr + KitchenAbvGr + Fireplaces + PavedDrive + WoodDeckSF + EnclosedPorch + ScreenPorch + PoolArea + MiscVal + SaleCondition + FirstFlrSF + SecondFlrSF, data = housesTrain)

plot(backwardResults)

plot(backwardModel)

results = housesTest
results$SalePrice = exp(predict.lm(backwardModel, newdata = housesTest))
results = results |>
  dplyr::select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Backward_Model.csv")
```

@@

# ols_step_both_aic does not have a direct output for the model determined so we have to write the results as we go and then copy the final model produced and paste it below
stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd, data = housesTrain)

plot(stepResults)

summary(stepModel)

plot(stepModel)

results = housesTest
results$SalePrice = exp(predict.lm(stepModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Step_Model.csv")
```

@@

results = housesTest
results$SalePrice = exp(predict(customModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export
write_csv(results, "Submissions/Custom_Model.csv")
```

>>>>

Define a function to determine the PRESS of each model
```{r}
PRESS = function(model) {
    i = residuals(model)/(1 - lm.influence(model)$hat)
    sum(i^2)
}
```

Forward step regression using olss
```{r}
# Remove high influence rows
# housesTrain2 = housesTrain |>
#   filter(!row_number() %in% c(121, 251, 272, 326, 584, 1004, 1231, 1276, 1299, 1322))

forwardResults = ols_step_forward_aic(lm(SalePrice ~ ., data = housesTrain), details = TRUE, progress = TRUE)
forwardModel = forwardResults$model
summary(forwardModel)

# Plot the stepping results impact on AIC
plot(forwardResults)

# Plot the residual / leverage functions
plot(forwardModel)

# Calculate the PRESS for the model
tibble(resids = residuals(forwardModel), hats = tidy(hatvalues(forwardModel))$x) |>
  filter(hats != 1) |>
  mutate(i = (resids / (1-hats))^2) |>
  select(i) |>
  pull() |>
  sum()

# Calculate values against the test data set
results = housesTest
results$SalePrice = exp(predict.lm(forwardModel, newdata = housesTest))
results = results |>
  dplyr::select(Id, SalePrice)

# Export and submit to Kaggle
write_csv(results, "Submissions/Forward_Model.csv")
```

@@

# Plot the stepping results impact on AIC
plot(backwardResults)

# Plot the residual / leverage functions
plot(backwardModel)

# Calculate the PRESS for the model
tibble(resids = residuals(backwardModel), hats = tidy(hatvalues(backwardModel))$x) |>
  filter(hats != 1) |>
  mutate(i = (resids / (1-hats))^2) |>
  select(i) |>
  pull() |>
  sum()

# Calculate values against the test data set
results = housesTest
results$SalePrice = exp(predict.lm(backwardModel, newdata = housesTest))
results = results |>
  dplyr::select(Id, SalePrice)

# Export and submit to Kaggle
write_csv(results, "Submissions/Backward_Model.csv")
```

@@

# ols_step_both_aic does not have a direct output for the model determined so we have to write the results as we go and then copy the final model produced and paste it below
stepModel = lm(SalePrice ~ OverallQual + GrLivArea + Neighborhood + OverallCond + YearBuilt + LotArea + RoofMatl + KitchenAbvGr + SaleCondition + Condition2 + Foundation + Fireplaces + Heating + ExterQual + Condition1 + PoolArea + ScreenPorch + WoodDeckSF + HeatingQC + CentralAir + BedroomAbvGr + FirstFlrSF + `2ndFlrSF` + Street + LandSlope + HalfBath + EnclosedPorch + SecondFlrSF + MiscVal + PavedDrive + BldgType + `1stFlrSF` + ExterCond + LotConfig + YearRemodAdd, data = housesTrain)
# Plot the stepping results impact on AIC
plot(stepResults)

# Plot the residual / leverage functions
plot(stepModel)

# Calculate the PRESS for the model
tibble(resids = residuals(stepModel), hats = tidy(hatvalues(stepModel))$x) |>
  filter(hats != 1) |>
  mutate(i = (resids / (1-hats))^2) |>
  select(i) |>
  pull() |>
  sum()

# Calculate values against the test data set
results = housesTest
results$SalePrice = exp(predict.lm(stepModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export and submit to Kaggle
write_csv(results, "Submissions/Step_Model.csv")
```

@@

# Calculate values against the test data setresults = housesTest
results = housesTest
results$SalePrice = exp(predict(customModel, newdata = housesTest))
results = results |>
  select(Id, SalePrice)

# Export and submit to Kaggle
write_csv(results, "Submissions/Custom_Model.csv")
```

### 2022-12-08 MSDS 6371 Project - Yule and Kondakindi.qmd
Minor updates
Writing updates

1.  Focusing on 3 specific neighborhoods in Ames, what is the best linear regression model that can be constructed solely utilizing the square footage of the living room and what can be interpreted from that model?
2.  Utilizing all available variables and focusing on every neighborhood, what are the best linear regression models that can be constructed?

@@

For this analysis, housing data was obtained between the years of 2006 and 2010 from 1,460 homes in the Ames area. Approximately 80 variables were collected for each home identifying various features that may contribute to a homes final sale price.

Upon analysis of the data, numerous variables were found to have a high amount of associated missing values and were subsequently removed from the analysis.

Many numerical variables in the available data showed high variance and were therefore given a log transformation to better perform in a linear regression model. The full list of variables that received a log transformation can be found below. It should be noted that sales price was among the variables.

@@

(Brief introduction to the questions of interest and the setting of the problem.)

@@

@@

@@

(Adj R2 Internal CV Press)

@@

For the second question of interest, the objective was to utilize all available data and variables to create the best linear regression model possible for predicting home sales prices. There are numerous techniques available for creating multi-linear regression models. This QOI focused on applying forward selection, backwards elimination, and stepwise selection to produce the highest performing models. Additionally, a final custom model was fit using cross validation techniques.

@@

The table below summarizes the key performance metrics associated with each of the three models produced. Each model was found to have the same Adjusted R Squared metric without rounding errors.

|     Model                 |  Adjusted  R2   |   CV Press    |   Kaggle Score    |
|---------------------------|-----------------|---------------|-------------------|
|   Forward Selection       | 0.9094          |   X           |   0.13496         |
|   Backward Elimination    | 0.9094          |   X           |   0.13554         |
|   Stepwise Selection      | 0.9094          |   X           |   0.13554         |

@@

(A short summary of the analysis)

>>>>

1.  Focusing on 3 specific neighborhoods in Ames, what is the best linear regression model that can be constructed solely utilizing the square footage of the living room and what can be interpreted from that model?
2.  Utilizing all available variables and focusing on every neighborhood, what are the best linear regression models that can be constructed? For this QOI, submissions will be judged based on a [Kaggle competition](https://www.kaggle.com/competitions/house-prices-advanced-regression-techniques), using produced model predictions to achieve the lowest root mean squared error.

@@

For this analysis, housing data from the Ames area was supplied between the years of 2006 and 2010 from 1,460 homes. Approximately 80 variables were collected for each home identifying various features that may contribute to a homes final sale price. The variables ranged in nature from quantitative values like square footage of certain parts of the home, to qualitative values like quality rankings and building materials.

Upon analysis of the data, numerous variables were found to have a high number of missing values and were subsequently removed from the analysis.

Many numerical variables in the available data demonstrated high variance. Log transformations were used to reduce the variance and ultimately improve the performance of linear regression models. The full list of variables that received a log transformation can be found below. It should be noted that sales price (the variable of interest) was among the variables recieving a log transformation.

@@

@@

For the first question of interest, the objective was to utilize only the greater living room square footage of a house to help predict the sales price in the three Ames neighborhoods of Brookside, Edwards, and North Ames.

Limiting the initial dataset to only these 3 neighborhoods results in 383 observations with 58, 100, and 225 for Brookside, Edwards, and North Ames respectively.

@@

Three models were fit to the data:

1. Average per neighborhood - A model which assumes only the average sales price for each neighborhood.
2. Varying y-intercept / same slopes - A model which allows for varying y-intercepts across each neighborhood while maintaining the same relationship in sales price and living room square footage.
3. Varying y-intercept / varying slopes - A model which allows for varying y-intercepts across each neighborhood in addition to varying relationships in sales price and living room square footage.

@@

The table below summarizes the key performance metrics associated with each of the three models produced. Overall the third model, which allows for varying y-intercepts and slopes across each neighborhood, outperforms the other 2 models.

|     Model                               |  Adjusted  R2   |   CV PRESS    |
|-----------------------------------------|-----------------|---------------|
|   Average per neighborhood              | 0.0880          |   26.44       |
|   Varying y-intercept / same slope      | 0.4857          |   15.00       |
|   Varying y-intercepts / varying slopes | 0.5056          |   14.61       |

@@

For the second question of interest, the objective was to utilize all available data and variables to create the best linear regression model possible for predicting home sales prices. There are numerous techniques available for creating multi-linear regression models. This QOI focused on applying forward selection, backwards elimination, and stepwise selection to produce the highest performing models.

@@

The table below summarizes the key performance metrics associated with each of the three models produced. Each model was found to have the same Adjusted R Squared metric within rounding errors. According to the predictive residual sum of squares metric, the stepwise selection model performed the best. When used to predict unknown home sales prices as part of the Kaggle competition, the Forward Selection model was found to achieve the best (lowest) Kaggle score of 0.13496. A custom model was excluded from this table as it was unable to achieve higher performance than the Forward and Stepwise selection methods.

|     Model                 |  Adjusted  R2   |   CV PRESS    |   Kaggle Score    |
|---------------------------|-----------------|---------------|-------------------|
|   Forward Selection       | 0.9094          |   26.12       |   0.13496         |
|   Backward Elimination    | 0.9094          |   25.95       |   0.13554         |
|   Stepwise Selection      | 0.9094          |   25.94       |   0.13554         |

@@

In this analysis, a data set of home sales prices from Ames, Iowa was explored using multi-linear regression techniques. A model was produced to predict the sales price of homes specifically in the North Ames, Edwards, and Brk Side neighborhoods using only the greater living area square footage. That model achieved an RSquared metric of XX, meaning XX% of the variance in home prices in those 3 neighborhoods can be explained soley by the living room squared footage.

Finally, 4 additional models were produced to predict homes sales prices based on all variables available. Various multi-linear regression stepping approaches were used and the forward selection method was found to perform the best overall with an adjusted RSquared metric of 0.9094 meaning over 90% of the variance in home prices could be explained by the variables and associated model. Additionally, a Kaggle competition score of 0.13496 was achieved by the model on a blind test.
